---
title: "bellabelt_Analysis"
author: "Amr Mohamed"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: False
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
```{r message=FALSE, warning=FALSE, include=FALSE}
# Data_prepration
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(janitor)
library(lubridate)
library(ggplot2)
library(readr)
```
```{r warning=FALSE, include=FALSE}
#Import files 
activity <- read.csv("C:/Users/Amr 10/Downloads/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
calories <- read.csv("C:/Users/Amr 10/Downloads/Fitabase Data 4.12.16-5.12.16/hourlyCalories_merged.csv")
intensities <- read.csv("C:/Users/Amr 10/Downloads/Fitabase Data 4.12.16-5.12.16/hourlyIntensities_merged.csv")
sleep <- read.csv("C:/Users/Amr 10/Downloads/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
weight <- read.csv("C:/Users/Amr 10/Downloads/Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
heart <- read.csv("C:/Users/Amr 10/Downloads/Fitabase Data 4.12.16-5.12.16/heartrate_seconds_merged.csv")

Sys.setlocale("LC_TIME","C")

```
```{r include=FALSE}
 # Activity date
activity$ActivityDate=as.POSIXct(activity$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
activity$date <- format(activity$ActivityDate, format = "%m/%d/%y")
activity$weekday <- format(activity$ActivityDate, format = "%A")
activity$weekday <- factor(activity$weekday, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
  
  # Calories date
calories$ActivityHour=as.POSIXct(calories$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
calories$time <- format(calories$ActivityHour, format = "%H:%M:%S")
calories$date <- format(calories$ActivityHour, format = "%m/%d/%y")

  # Intensities date
intensities$ActivityHour=as.POSIXct(intensities$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
intensities$time <- format(intensities$ActivityHour, format = "%H:%M:%S")
intensities$date <- format(intensities$ActivityHour, format = "%m/%d/%y")

  # sleep 
sleep$SleepDay=as.POSIXct(sleep$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
sleep$date <- format(sleep$SleepDay, format = "%m/%d/%y")

  # weight 
weight$Date_Time=as.POSIXct(weight$Date, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
weight$Date <- format(weight$Date_Time, format = "%m/%d/%y")
weight$time <- format(weight$Date_Time, format = "%H:%M:%S")
  # Heart 
heart$Date_Time=as.POSIXct(heart$Time, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
heart$Date <- format(heart$Date_Time, format = "%m/%d/%y")
heart$time <- format(heart$Date_Time, format = "%H:%M:%S")

```

##  BUSINESS TASK: 

### Analyze Fitbit data to gain insight and help guide marketing strategy for Bellabeat to grow as a global player.

## Stakeholders: 
### Bellabeat marketing analytics team

```{r warning=FALSE, include=FALSE}
#Remove duplicates 

  #Activity 
activity <- activity %>%
  distinct() %>%
  clean_names() %>%
  drop_na()

  #Calories
calories <- calories %>%
  distinct() %>%
  clean_names() %>%
  drop_na()

  #Intensities
intensities <- intensities %>%
  distinct() %>%
  clean_names() %>%
  drop_na() 
  
  #sleep 
sleep <- sleep %>%
  distinct() %>%
  clean_names() %>%
  drop_na()

  #weight 
weight <- weight %>% 
  distinct() %>%
  clean_names() %>%
  drop_na()

  #heart 
heart <- heart %>% 
  distinct() %>%
  clean_names() %>%
  drop_na()

```

```{r include=FALSE}
#Merge activity with weight and Visualize 
activity_with_weight <- merge(activity,weight, by= 'id')
```
# This summary of Activity&Weight dataset
```{r echo=FALSE}
activity_with_weight %>% 
  select(total_steps,weekday,weight_kg,fat,bmi,calories) %>%
  summary()
```
## Visualize Losing Calories with total_Steps
```{r echo=FALSE}
  ggplot(data = activity_with_weight,aes(total_steps,calories)) + 
  geom_point() + geom_smooth() + labs(title="Total Steps vs. Calories")
```

__Note There's a positive correlation between total steps and calories__



# Visualize Most days people be active
```{r echo=FALSE}
  ggplot(data = activity_with_weight, aes(weekday,total_steps)) +
    geom_col() +
    geom_text(aes(label = total_steps), vjust = 0, colour = "white",check_overlap = TRUE) +
    labs(title="total_steps vs. weekday")
```

__Note that most steps active at Mid-Week__

```{r include=FALSE}
#merge activity with sleep and visualize
  activity_with_sleep <- merge(activity,sleep, by=c("id", "date"))
```


# Visualize  sedentary_minutes Active per minuites sleep 
```{r echo=FALSE}
    ggplot(data =activity_with_sleep,aes(total_minutes_asleep, sedentary_minutes)) +
    geom_point() + geom_smooth()+  labs(title="Minutes Asleep vs. Sedentary Minutes")
```

__Note that There's a negative correlation__ 

```{r echo=FALSE}
  ggplot(data =activity_with_sleep,aes(total_minutes_asleep, calories))+
    geom_point() + geom_smooth()+  labs(title="Minutes Asleep vs. losingcalories")

```


# Recomendations 

  __1. Bellabeat can encourage people to take at least 8 000 explaining the         benefits for their health__

  __2.If users want to lose weight, it’s probably a good idea to control daily        calorie consumption. Bellabeat can suggest some ideas for low-calorie lunch and dinner.__

  __3. If users want to improve their sleep, Bellabeat should consider using app    notifications to go to bed.__

  __4. if users want to improve their sleep, the Bellabeat app can recommend        reducing sedentary time.__

  __5. Can send notifications to the audience to make more workout at the first     of week.__ 